---
output: html_document
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Spatial Interpolation: meuse

## Data

```{r, warning=FALSE, message=FALSE}
library(sp)
library(gstat)
library(sf)
library(mapview)
library(tidyverse)
```

```{r}
data(meuse)
head(meuse)
ggplot(data = meuse) + geom_point(aes(x, y))
```

```{r}
data(meuse.grid)
head(meuse.grid)
ggplot(data = meuse.grid) + geom_point(aes(x, y))
```





## Summary Statistics

```{r}
summary(meuse)
```

### Bubble plot

```{r}
ggplot(data = meuse) + geom_point(aes(x, y, size=zinc))
```

## Spatial Data Wrangling

```{r}
class(meuse)
class(meuse.grid)
```

### Convert the meuse and meuse.grid data frames to sf objects using the st_as_sf()

```{r}
meuse2 <- st_as_sf(meuse, coords = c("x", "y"), crs = 28992)
meuse.grid2 <- st_as_sf(meuse.grid, coords = c("x", "y"),
                       crs = 28992)
```

### Recheck class type

```{r}
class(meuse2)
class(meuse.grid2)
```

```{r}
head(meuse2)
head(meuse.grid2)
```

### Interactive maps 

```{r}
mapview(meuse2, zcol = "zinc",  map.types = "CartoDB.Voyager")
mapview(meuse.grid2,  map.types = "CartoDB.Voyager")
```

## EDA

```{r}
ggplot(meuse, aes(x=dist, y=zinc)) + geom_point()
```

```{r}
ggplot(meuse, aes(x=dist, y=log(zinc))) + geom_point()
```

```{r}
ggplot(meuse, aes(x=sqrt(dist), y=log(zinc))) + geom_point()
```

```{r, warning=FALSE, message=FALSE}
library(GGally)
ggpairs(meuse)
```

```{r}
ggpairs(meuse[,3:9])


```

### Simple Linear Regression Model

```{r}
lm <- lm(log(zinc)~sqrt(dist), meuse)
meuse$residuals <- residuals(lm)
meuse$fitted <- fitted(lm) 
meuse$fitted2 <- predict(lm, meuse) 
meuse$fitted.s <- predict(lm, meuse) - mean(predict(lm, meuse))
head(meuse)
```

```{r}
library(viridis)
ggplot(meuse, aes(x=x, y=y, col=fitted))+geom_point()+scale_color_viridis()
ggplot(meuse, aes(x=x, y=y, col=residuals))+geom_point()+scale_color_viridis()
ggplot(meuse, aes(x=x, y=y, col=fitted.s))+geom_point()+scale_color_viridis()
```

## Inverse Distance Interpolation

```{r}
library(stars) |> suppressPackageStartupMessages()
library(gstat)
idw_result <- idw(zinc~1, meuse2, meuse.grid2)
idw_result
# Convert IDW output to a data frame
idw_df <- as.data.frame(idw_result)
head(idw_df)
colnames(idw_df) <- c("x", "y", "zinc_pred")
head(idw_df)

```

### Visualize IDW values

```{r}
ggplot(data = meuse2) + geom_sf(data = idw_result, aes(color = var1.pred)) +
  geom_sf() +
  scale_color_viridis() + theme_bw()
```

## Spatial Kriging

### Lagged Scatter plots

```{r}
hscat(log(zinc)~1, meuse2, (0:9)*100)
```

### Variogram Cloud

```{r}
vc <- variogram(log(zinc) ~ 1, meuse2, cloud = TRUE)
plot(vc)
```

### Binned Variogram/ Sample Variogram

```{r}
v <- variogram(log(zinc) ~ 1, data = meuse2)
plot(v)
```

### List of available models for variograms

```{r}
vgm()
show.vgms(par.strip.text = list(cex = 0.75))
```

### Fitting Variogram Models

### Trial 1

#### Assess our initial model

```{r}
vg.fit1 <- vgm(psill = 0.5, model = "Sph",
                range = 900, nugget = 0.1)
vg.fit1 
plot(v, vg.fit1 , cutoff = 1000, cex = 1.5)
```

### Fit a variogram providing the sample variogram

```{r}
fv <- fit.variogram(object = v,
                    model = vgm(psill = 0.5, model = "Sph",
                                range = 900, nugget = 0.1))
fv
attr(fv, "SSErr")
plot(v, fv, cex = 1.5)
```

### Trial 2


```{r}
fv2 <- fit.variogram(object = v,
                    model = vgm(psill = 0.6, model = "Sph",range = 900, nugget = 0.1))
fv2
attr(fv2, "SSErr")
plot(v, fv2, cex = 1.5)
```

### Trial 3


```{r}
fv3 <- fit.variogram(object = v,
                    model = vgm(psill = 1, model = "Exp",range = 800, nugget = 0.1))
fv3
attr(fv3, "SSErr")
plot(v, fv3, cex = 1.5)
```

#### If the process is isotropic we can directly go for kriging

```{r}
library(ggplot2)
library(viridis)

k <- gstat(formula = log(zinc) ~ 1, data = meuse2, model = fv)
kpred <- predict(k, meuse.grid2)
head(kpred)
```

#### Visualisation of Predictions

```{r}
ggplot() + geom_sf(data = kpred, aes(color = var1.pred)) +
  geom_sf(data = meuse2) +
  scale_color_viridis(name = "log(zinc)") + theme_bw()

ggplot() + geom_sf(data = kpred, aes(color = var1.var)) +
  geom_sf(data = meuse2) +
  scale_color_viridis(name = "variance") + theme_bw()
```

## Detecting Anisotropy

```{r}
vg.dir <- variogram(log(zinc)~1, alpha = c(0, 45, 90, 135), meuse2)
plot(vg.dir)
```

## Fit Anisotropic Variogram Models

The function call vgm(psill=0.59, model = "Sph", range = 1200, nugget = 0.05, anis = c(45, 0.4)) in R creates a variogram model using the vgm() function from the gstat package. Here's what each parameter means:

- psill = 0.59: This is the partial sill, representing the difference between the sill (total variance) and the nugget effect. It defines the maximum semivariance minus the nugget.

- model = "Sph": The model type is Spherical, one of the commonly used variogram models. It describes how spatial correlation decreases with distance.

- range = 1200: This is the range, the distance at which the spatial correlation drops to near zero (or 95% of the sill for the spherical model). Beyond this distance, observations are considered uncorrelated.

- nugget = 0.05: The nugget effect, representing spatial variability at very small scales or measurement errors. It accounts for discontinuity at very short distances.

- anis = c(45, 0.4): Specifies anisotropy, meaning the spatial correlation differs by direction.

- 45 indicates the direction (angle in degrees from the x-axis) where the correlation is strongest.

- 0.4 is the anisotropy ratio, meaning correlation decays faster in the perpendicular direction. A value of 0.4 means that in the perpendicular direction, the range is only 40% of the main direction's range.

```{r}
fit.ani <- vgm(psill=0.59, model = "Sph", range = 1200, nugget = 0.05, anis = c(45, 0.4))
plot(vg.dir, fit.ani)
```

This variogram model describes spatial dependence using a spherical variogram with anisotropy. The correlation extends up to 1200 units in the primary direction (45°), but only 480 units (1200 × 0.4) in the perpendicular direction. There is a small nugget effect (0.05), indicating slight short-scale variation or measurement noise.

## Variogram Maps

Another way to detect directional dependence

```{r}
vg.map <- variogram(log(zinc)~1, meuse2, map=T, cutoff=1000, width=100)
plot(vg.map, threshold=5)
```


#### Explanation of the Code: `vg.map <- variogram(log(zinc) ~ 1, meuse, map = TRUE, cutoff = 1000, width = 100)`



This code computes and visualizes the variogram map of the log-transformed zinc concentrations in the Meuse dataset.

1. This function calculates a directional semivariogram for the spatial data.

- variogram(log(zinc) ~ 1, meuse, ...)

      Computes the experimental semivariogram for log-transformed zinc concentrations. The ~ 1 formula means it calculates the variogram without considering covariates (i.e., only based on spatial distances).

- map = TRUE

    Instead of returning a traditional one-dimensional variogram, this creates a spatial variogram map, which helps identify anisotropy (directional dependence of spatial variation).

- cutoff = 1000

    Defines the maximum lag distance (1,000 meters). Pairs of points separated by distances greater than this are ignored.

- width = 100

    Sets the bin width for grouping distances. Distance intervals of 100 meters are used when computing the variogram.

2. Breakdown of plot(vg.map, threshold = 5)
This function plots the variogram map.

- plot(vg.map, ...)

    Plots the spatial variogram map, where each pixel represents a directional semivariance value for a given spatial lag.

- threshold = 5

    Sets a threshold to filter high semivariance values to make the plot more interpretable.
    
## Model residuals obtain from `log(zinc)~sqrt(dist)`


```{r}
vgreg.est <- variogram(log(zinc)~sqrt(dist), meuse2)
plot(vgreg.est)
vgreg.fit <- fit.variogram(vgreg.est, vgm(0.2, "Sph", 800, 0.05))
vgreg.fit
plot(vgreg.est, vgreg.fit, cex = 1.5)
```

```{r}
vgreg.dir <- variogram(log(zinc)~sqrt(dist), meuse2, alpha=c(0, 45, 90, 135))
plot(vgreg.dir)
```

```{r}
plot(vgreg.dir, vgreg.fit, cex = 1.5)
```

The directional dependence is much less obvious in this case.

Another way to show this is variogram maps

```{r}
vgreg.map <- variogram(log(zinc)~sqrt(dist), meuse2, map=T, cutoff=1000, width=100)
plot(vgreg.map, threshold=5)
```

## Simple Kriging
 